Active Learning for SAT-Solver Benchmarking

### 2022-10-06 paper/paper.tex
Replace tables in evaluation with plots

\begin{table}[tb!]
  %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
  \centering
  \caption{
  	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
  	Rows with no changes in comparison to their predecessor are hidden.
  	%JB: no change in what? all columns except delta?
  	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
  	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
  	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
  	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
  }
  \label{tab:paretotable}
  \vspace{0.2cm}
  \begin{tabular}{
    >{\centering\arraybackslash}m{0.6cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.6cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
    >{\centering\arraybackslash}m{1.5cm}
  }
    \hline
    $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
    \hline
    0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
    0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
    0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
    0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
    0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
    0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
    0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
    $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
    0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
    0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
    \hline
    0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
    \hline
    0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
    0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
    1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
    \hline
  \end{tabular}
\end{table}

@@

\begin{table}[tb!]
  %JB: "fraction" here only draws family-specific comparison of sample vs. full dataset, but does not compare overall distribution of families in sample/dataset; might be interesting to have scatterplot with "(relative) frequency of family in sample" vs. "(relative) frequency of family in dataset" (does not allow to show all family labels (though we probably could manually label some families), but allows to show all families)
  %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
  \centering
  \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
  \label{tab:families}
  \begin{subtable}[t]{0.48\textwidth}
    \centering
    \caption{Sorted by the average number of occurrences in the sample}
    \label{tab:familiesa}

    \begin{tabular}{
      >{\centering\arraybackslash}m{1.8cm}
      >{\centering\arraybackslash}m{1.2cm}
      >{\centering\arraybackslash}m{1.2cm}
      >{\centering\arraybackslash}m{1.2cm}
    }
      \hline
      Family & In Sample & In Dataset & Fraction (\%) \\
      \hline
      planning              & 19.14 &  333 &  5.7 \\
      subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
      crypto-graphy         & 13.11 &  311 &  4.2 \\
      pigeon-hole           &  7.55 &  151 &  5.0 \\
      fpga-routing          &  6.16 &  73  &  8.4 \\
      rbsat                 &  5.93 &  113 &  5.2 \\
      quasigroup-completion &  5.23 &  210 &  2.5 \\
      unknown               &  4.93 &  169 &  2.9 \\
      miter                 &  4.84 &  199 &  2.4 \\
      bitvector             &  4.73 &  214 &  2.2 \\
      % clique-coloring       &  4.21 &  64  &  0.066 \\
      % mosoi-289             &  3.75 &  31  &  0.121 \\
      % hardware-verification &  3.70 &  344 &  0.011 \\
      % hardware-bmc          &  2.77 &  78  &  0.035 \\
      % graph-based           &  2.77 &  134 &  0.021 \\
      % coloring              &  2.66 &  41  &  0.065 \\
      % waerden               &  2.55 &  43  &  0.059 \\
      % graph-coloring        &  2.50 &  87  &  0.029 \\
      % auto-correlation      &  2.39 &  46  &  0.052 \\
      % knights-problem       &  2.25 &  17  &  0.132 \\
      %\hline
      Remaining families    & 104.45 & 3407 & 3.1 \\
      \hline
      Sum                   & 192.34 & 5355 & 3.6 \\
      \hline
    \end{tabular}
  \end{subtable}
  \hfill
  \begin{subtable}[t]{0.48\textwidth}
    \centering
    \caption{Sorted by the fraction of sampled occurrences}
    \label{tab:familiesb}

    \begin{tabular}{
      >{\centering\arraybackslash}m{1.8cm}
      >{\centering\arraybackslash}m{1.2cm}
      >{\centering\arraybackslash}m{1.2cm}
      >{\centering\arraybackslash}m{1.2cm}
    }
      \hline
      Family & In Sample & In Dataset & Fraction (\%) \\
      \hline
      circuit-multiplier          & 1.59    &      9      &     17.7 \\
      subset-cardinality          & 1.55    &      9      &     17.3 \\
      ssp-0                       & 0.34    &      2      &     17.0 \\
      sliding-puzzle              & 1.52    &      11     &     13.8 \\
      knights-problem             & 2.25    &      17     &     13.2 \\
      genurq                      & 1.55    &      12     &     12.9 \\
      stone                       & 1.39    &      11     &     12.7 \\
      mosoi-289                   & 3.75    &      31     &     12.1 \\
      anti-bandwidth              & 1.52    &      13     &     11.7 \\
      random                      & 0.91    &      8      &     11.4 \\
      % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
      % social-golfer               & 0.95    &      9      &     0.105159 \\
      % product-configuration       & 0.52    &      5      &     0.103571 \\
      % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
      % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
      % sgen-balanced               & 0.27    &      3      &     0.089286 \\
      % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
      % fpga-routing                & 6.16    &      73     &     0.084393 \\
      % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
      % core-based-generator        & 1.00    &      13     &     0.076923 \\
      %\hline
      Remaining families    & 175.97 & 5232 & 3.4 \\
      \hline
      Sum                   & 192.34 & 5355 & 3.6 \\
      \hline
    \end{tabular}
  \end{subtable}
\end{table}

%\vspace{10cm}
\section{Conclusion}
In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
We also took a brief look at which instance families are the most prevalent when sampling instances.

>>>>

% \begin{table}[tb!]
%   %JB: really hard to extract take-aways by looking at this table (several hyperprameter settings and several metrics, the hyperparameter settings of our approach change from row to row)
%   \centering
%   \caption{
%   	Pareto-optimal approaches for different choices of $\delta$ regarding our smaller hyper-parameter tuning dataset.
%   	Rows with no changes in comparison to their predecessor are hidden.
%   	%JB: no change in what? all columns except delta?
%   	Ranking approaches are either predicted cluster-label ranking (Pred.) or partially observed PAR-2 ranking (PAR-2).
%   	Selection approaches are either uncertainty-based selection (Uncert.) or information-gain-based selection (IG).
%   	Stopping criteria are either ranking convergence (Conv.) or Wilcoxon-based (Wilcoxon).
%   	%JB: How are the values for ranking/selection/stopping determined -- whatever is best for a particular delta?
%   }
%   \label{tab:paretotable}
%   \vspace{0.2cm}
%   \begin{tabular}{
%     >{\centering\arraybackslash}m{0.6cm}
%     >{\centering\arraybackslash}m{1.6cm}
%     >{\centering\arraybackslash}m{1.6cm}
%     >{\centering\arraybackslash}m{1.6cm}
%     >{\centering\arraybackslash}m{1.5cm}
%     >{\centering\arraybackslash}m{1.5cm}
%     >{\centering\arraybackslash}m{1.5cm}
%     >{\centering\arraybackslash}m{1.5cm}
%   }
%     \hline
%     $\delta$ & Ranking & Selection & Stopping & Fraction Runtime ($O_{\operatorname{rt}}$) & Fraction Instances & $O_{\operatorname{acc}}$ & $O_{\delta}$ \\
%     \hline
%     0.00 & PAR-2 & Uncert. & Conv. & 0.021 & 0.025 & 0.593 & 0.979 \\
%     0.05 & Pred. & Uncert. & Conv. & 0.024 & 0.158 & 0.765 & 0.966 \\
%     0.10 & Pred. & Uncert. & Conv. & 0.025 & 0.128 & 0.779 & 0.955 \\
%     0.15 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.946 \\
%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
%     0.50 & Pred. & IG & Conv. & 0.028 & 0.112 & 0.802 & 0.887 \\
%     0.55 & Pred. & Uncert. & Conv. & 0.053 & 0.142 & 0.824 & 0.880 \\
%     0.60 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.880 \\
%     $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ & $\vdots$ \\
%     0.75 & Pred. & Uncert. & Conv. & 0.146 & 0.082 & 0.897 & 0.886 \\
%     0.80 & Pred. & Uncert. & Conv. & 0.219 & 0.128 & 0.917 & 0.890 \\
%     \hline
%     0.85 & Pred. & Uncert. & Conv. & 0.272 & 0.171 & 0.927 & 0.897 \\
%     \hline
%     0.90 & Pred. & Uncert. & Wilcoxon & 0.894 & 0.888 & 0.996 & 0.907 \\
%     0.95 & Pred. & Uncert. & Wilcoxon & 0.931 & 0.924 & 1.000 & 0.953 \\
%     1.00 & Pred. & Uncert. & Wilcoxon & 0.989 & 0.981 & 1.000 & 1.000 \\
%     \hline
%   \end{tabular}
% \end{table}

\begin{figure}[tb!]
  \centering
  \resizebox{0.78\textwidth}{!}{
    \graphicspath{{../plots/}}
    \input{../plots/anni_train_optimization_goal.pgf}
  }
  \caption{
    Scatter plot showing different instantiations of our active-learning strategy for different $\delta$ on our smaller hyper-parameter-tuning dataset.
    The x-axis shows the fraction of runtime $O_{\operatorname{rt}}$ that a particular instantiation needs.
    The y-axis shows the fraction of instances that a particular instantiation selects to arrive at its conclusion.
    The color indicates the weighting between different optimization goals $\delta \in \left[0, 1\right]$.
    A small $\delta$ refers to favoring runtime over accuracy and a big $\delta$ refers to favoring accuracy over runtime.
  }
  \label{fig:annitrainoptgoal}
\end{figure}

@@

% \begin{table}[tb!]
%   %JB: "fraction" here only draws family-specific comparison of sample vs. full dataset, but does not compare overall distribution of families in sample/dataset; might be interesting to have scatterplot with "(relative) frequency of family in sample" vs. "(relative) frequency of family in dataset" (does not allow to show all family labels (though we probably could manually label some families), but allows to show all families)
%   %JB: could also correlate "(relative) frequency of family in sample" with "(relative) frequency of family in dataset"
%   \centering
%   \caption{Top-10 most \emph{important} families regarding the average occurrences and the fraction of selected occurrences regarding our AL approach}
%   \label{tab:families}
%   \begin{subtable}[t]{0.48\textwidth}
%     \centering
%     \caption{Sorted by the average number of occurrences in the sample}
%     \label{tab:familiesa}

%     \begin{tabular}{
%       >{\centering\arraybackslash}m{1.8cm}
%       >{\centering\arraybackslash}m{1.2cm}
%       >{\centering\arraybackslash}m{1.2cm}
%       >{\centering\arraybackslash}m{1.2cm}
%     }
%       \hline
%       Family & In Sample & In Dataset & Fraction (\%) \\
%       \hline
%       planning              & 19.14 &  333 &  5.7 \\
%       subgraph-isomorphism  & 16.27 &  175 &  9.3 \\
%       crypto-graphy         & 13.11 &  311 &  4.2 \\
%       pigeon-hole           &  7.55 &  151 &  5.0 \\
%       fpga-routing          &  6.16 &  73  &  8.4 \\
%       rbsat                 &  5.93 &  113 &  5.2 \\
%       quasigroup-completion &  5.23 &  210 &  2.5 \\
%       unknown               &  4.93 &  169 &  2.9 \\
%       miter                 &  4.84 &  199 &  2.4 \\
%       bitvector             &  4.73 &  214 &  2.2 \\
%       % clique-coloring       &  4.21 &  64  &  0.066 \\
%       % mosoi-289             &  3.75 &  31  &  0.121 \\
%       % hardware-verification &  3.70 &  344 &  0.011 \\
%       % hardware-bmc          &  2.77 &  78  &  0.035 \\
%       % graph-based           &  2.77 &  134 &  0.021 \\
%       % coloring              &  2.66 &  41  &  0.065 \\
%       % waerden               &  2.55 &  43  &  0.059 \\
%       % graph-coloring        &  2.50 &  87  &  0.029 \\
%       % auto-correlation      &  2.39 &  46  &  0.052 \\
%       % knights-problem       &  2.25 &  17  &  0.132 \\
%       %\hline
%       Remaining families    & 104.45 & 3407 & 3.1 \\
%       \hline
%       Sum                   & 192.34 & 5355 & 3.6 \\
%       \hline
%     \end{tabular}
%   \end{subtable}
%   \hfill
%   \begin{subtable}[t]{0.48\textwidth}
%     \centering
%     \caption{Sorted by the fraction of sampled occurrences}
%     \label{tab:familiesb}

%     \begin{tabular}{
%       >{\centering\arraybackslash}m{1.8cm}
%       >{\centering\arraybackslash}m{1.2cm}
%       >{\centering\arraybackslash}m{1.2cm}
%       >{\centering\arraybackslash}m{1.2cm}
%     }
%       \hline
%       Family & In Sample & In Dataset & Fraction (\%) \\
%       \hline
%       circuit-multiplier          & 1.59    &      9      &     17.7 \\
%       subset-cardinality          & 1.55    &      9      &     17.3 \\
%       ssp-0                       & 0.34    &      2      &     17.0 \\
%       sliding-puzzle              & 1.52    &      11     &     13.8 \\
%       knights-problem             & 2.25    &      17     &     13.2 \\
%       genurq                      & 1.55    &      12     &     12.9 \\
%       stone                       & 1.39    &      11     &     12.7 \\
%       mosoi-289                   & 3.75    &      31     &     12.1 \\
%       anti-bandwidth              & 1.52    &      13     &     11.7 \\
%       random                      & 0.91    &      8      &     11.4 \\
%       % spectrum-repacking          & 1.30    &      12     &     0.108631 \\
%       % social-golfer               & 0.95    &      9      &     0.105159 \\
%       % product-configuration       & 0.52    &      5      &     0.103571 \\
%       % minimal-disagreement-parity & 1.54    &      16     &     0.095982 \\
%       % subgraph-isomorphism        & 16.27   &       175   &     0.092959 \\
%       % sgen-balanced               & 0.27    &      3      &     0.089286 \\
%       % pythagorean-triples         & 1.82    &      21     &     0.086735 \\
%       % fpga-routing                & 6.16    &      73     &     0.084393 \\
%       % petrinet-concurrency        & 2.11    &      25     &     0.084286 \\
%       % core-based-generator        & 1.00    &      13     &     0.076923 \\
%       %\hline
%       Remaining families    & 175.97 & 5232 & 3.4 \\
%       \hline
%       Sum                   & 192.34 & 5355 & 3.6 \\
%       \hline
%     \end{tabular}
%   \end{subtable}
% \end{table}

\begin{figure}[tb!]
  \centering
  \resizebox{0.85\textwidth}{!}{
    \graphicspath{{../plots/}}
    \input{../plots/anni_final_families.pgf}
  }
  \caption{
    Scatter plot showing the \emph{importance} of different instance families.
    The x-axis shows the fraction of instances within a particular family in the complete SAT Competition~2022 Anniversary Track benchmark set~\cite{sat2022}.
    The y-axis shows the average fraction of instances within a particular family that is selected by our active-learning strategies.
    The dashed line represents families that are represented with the same fraction in, both, our dataset and the incremental benchmark sets $\tilde{\mathcal{I}}$ that are chosen by our approach.
  }
  \label{fig:annifinalfamilies}
\end{figure}

\section{Conclusion}
In this work, we have discussed possible solutions to the \emph{New-Solver Problem}:
Given a new solver, we want to find its ranking amidst its competitors.
Our approaches provide accurate ranking predictions while only needing a fraction of the runtime resources that a full evaluation on all benchmarking instances would need.
We use a runtime discretization technique as this enables us to transform the regression problem of directly predicting runtimes into the much simpler notion of classification.
We have shown that, albeit being more simple, the classification of discrete runtime labels produces good results.
We have evaluated several ranking algorithms, instance-selection approaches, and stopping criteria within our sequential active-learning process.
A model-uncertainty-based selection approach in combination with runtime-label-prediction ranking and a ranking-convergence stopping criterion showed the consistently best results.
%JB: "consistently best" -- it's the best approach for one delta, but not sure regarding other deltas nor its sensitivity to hyperparameters
We also took a brief look at which instance families are the most prevalent when sampling instances.
